---
title: "Assignment 4"
author: "Gemma Holt"
date: "10/262020"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE)
```

## Setup
```{r}
library(osmdata)
library(opentripplanner)
library(tidyverse)
library(sf)
library(ggthemes)
library(ggspatial)
```


## Load locations
I am examining the location of public libraries in Vancouver, BC.
```{r}
van_libraries <- st_read("https://opendata.vancouver.ca/explore/dataset/libraries/download/?format=kml&timezone=America/New_York&lang=en", quiet = TRUE)
```


## Get street data

```{r}
van_plane <- "+proj=aea +lat_1=50 +lat_2=58.5 +lat_0=45 +lon_0=-126 +x_0=1000000 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"
## Is this the right projection?

vancouver_street_query <- opq(bbox = 'Vancouver BC Canada') %>%
  add_osm_feature(key = 'highway')

vancouver_street_query %>%
   osmdata_xml(file = 'OTP/graphs/default/vancouver_streets.osm')

vancouver_street_features <- vancouver_street_query %>%
  osmdata_sf()

vancouver_streets <- vancouver_street_features$osm_lines %>%
  st_transform(crs=van_plane)
```


## Plotting the streets
```{r}
ggplot(vancouver_streets) +
  geom_sf() +
  theme_map()
```


## Launch Open Trip Planner
```{r, message=FALSE, results="hide"}
path_data <- file.path(getwd(), "OTP")
path_otp <- paste(path_data, "otp.jar",sep = "/")
otp_build_graph(otp = path_otp, dir = path_data, memory = 1024) 
```

```{r}
otp_setup(otp = path_otp, dir = path_data, memory =1024)
otpcon <- otp_connect()
```


## Creating isochrones
```{r}
iso_10min_walk <- 
  otp_isochrone(otpcon = otpcon, fromPlace = van_libraries,
                mode = "WALK", cutoffSec = 600) %>%
  st_transform(crs = van_plane) %>%
  mutate(mode = "walk")

iso_10min_drive <- 
  otp_isochrone(otpcon = otpcon, fromPlace = van_libraries, 
                mode = "CAR", cutoffSec = 600) %>%
  st_transform(crs = van_plane) %>%
  mutate(mode = "drive")

iso_10min_bike <- 
  otp_isochrone(otpcon = otpcon, fromPlace = van_libraries, 
                mode = "BICYCLE", cutoffSec = 600) %>%
  st_transform(crs = van_plane) %>%
  mutate(mode = "bike")

iso_all_modes <- rbind(iso_10min_drive, iso_10min_walk,iso_10min_bike)

iso_walk_bike <- rbind(iso_10min_walk, iso_10min_bike)
```


## The area of a ten-minute walkshed
```{r}
right_side <- st_bbox(iso_10min_walk)$xmax
left_side  <- st_bbox(iso_10min_walk)$xmin
top_side <- st_bbox(iso_10min_walk)$ymax
bottom_side <- st_bbox(iso_10min_walk)$ymin

ggplot(iso_10min_walk) +
  annotation_map_tile(zoomin = 0, type = "cartolight", progress = "none") +
  geom_sf(aes(fill = mode), alpha = 0.5) +
  geom_sf(data = van_libraries) +
  coord_sf(xlim = c(left_side, right_side), 
           ylim = c(bottom_side, top_side), expand = FALSE) +
  scale_fill_viridis_d(name = "Area that is reachable within 10 minutes",
                       labels = c("By foot")) +
  theme_map() +
  labs(caption = "Basemap Copyright OpenStreetMap contributors")
```

## The area of a ten-minute walk and bikeshed
```{r}
right_side <- st_bbox(iso_walk_bike)$xmax
left_side  <- st_bbox(iso_walk_bike)$xmin
top_side <- st_bbox(iso_walk_bike)$ymax
bottom_side <- st_bbox(iso_walk_bike)$ymin

ggplot(iso_walk_bike) +
  annotation_map_tile(zoomin = 0, type = "cartolight", progress = "none") +
  geom_sf(aes(fill = mode), alpha = 0.5) +
  geom_sf(data = van_libraries) +
  coord_sf(xlim = c(left_side, right_side), 
           ylim = c(bottom_side, top_side), expand = FALSE) +
  scale_fill_viridis_d(name = "Area that is reachable within 10 minutes",
                       labels = c("By bike", "By foot")) +
  theme_map() +
  labs(caption = "Basemap Copyright OpenStreetMap contributors")
```


## Comparing walksheds to bikesheds
There was a duplicate value that was inhibiting the scatter plot, so I first removed that from the table. 
```{r}
duplicated(iso_walk_bike)
iso_walk_bike <- iso_walk_bike[!duplicated(iso_walk_bike), ]

iso_areas <- iso_walk_bike %>%
  mutate(area = st_area(iso_walk_bike)) %>%
  st_set_geometry(NULL) %>%
  pivot_wider(names_from = mode, values_from = area)

ggplot(iso_areas, 
  aes(x = as.numeric(walk), y = as.numeric(bike))) +
  geom_point() +
  scale_x_continuous(name = 
            "Area within a ten-minute walking distance\nof a public library\n(square km)",
            breaks = breaks <- seq(100000, 1300000, by = 200000),
            labels = breaks / 1000000) +
  scale_y_continuous(name = 
            "Area within a ten-minute biking distance\nof a public library\n(square km)",
            breaks = breaks <- seq(0, 10000000, by = 2000000),
            labels = breaks / 10000000) +
  theme_bw()
```


## Schools within a ten-minute walk of a library
First I filtered the data to only display schools that are within a ten-minute walk of a public library. These were two variables that I looked at last week (in Boston). I was interested in seeing how the data might look similar or different in Vancouver. 
```{r}
van_schools <- st_read("https://opendata.vancouver.ca/explore/dataset/schools/download/?format=kml&timezone=America/New_York&lang=en", quiet = TRUE)
```


```{r}
van_schools <- van_schools %>% 
  st_transform(van_plane)

walk_schools <- van_schools[iso_10min_walk, ]

ggplot(iso_10min_walk) +
  annotation_map_tile(zoomin = 0, type = "cartolight", progress = "none") +
  geom_sf() +
  geom_sf(data = walk_schools, 
          color = "orange", 
          size = 2) +
  theme_map()

```

## Percentage of schools that are not within a ten-minute walk to a library
```{r}
van_schools <- van_schools %>%
  st_join(walk_schools) %>%
  mutate(by_library = !is.na(Name.y))

n_walk_schools <- sum(van_schools$by_library)
n_walk_schools

```

```{r}
n_van_schools <- length(van_schools$by_library)
pct_walk_schools <- n_walk_schools / n_van_schools
pct_walk_schools
```
```{r}
left_side  <- st_bbox(iso_all_modes)$xmin
top_side <- st_bbox(iso_all_modes)$ymax

ggplot(iso_10min_walk) +
  annotation_map_tile(zoomin = 0, type = "cartolight", progress = "none") +
  geom_sf(fill = "grey92", color = "grey46") +
  geom_sf(data = van_schools, size = 2,
          aes(color = by_library)) +
  scale_color_manual(values = c("lightblue", "orange"),
  name = "Schools\nby distance to a library", 
  labels = c("No library within 10 minute walk",
                     "Library within 10 minute walk")) +
  annotation_scale(location = "br") +
  annotate(geom = "text", x = left_side, 
           y = top_side, 
           label = paste("Of the ",
                   prettyNum(n_van_schools, big.mark = ","),
                   " schools in Vancouver\n",
                   prettyNum(n_walk_schools, big.mark = ","),
                   " (",
                   prettyNum(100*pct_walk_schools, digits = 0),
                   "%) are within a 10-minute\nwalk of a library.",
                   sep = ""),
           hjust = -0.01, vjust = .9, size = 3) +
  theme_map() +
  ggtitle("Vancouver schools within a ten-minute walk to libraries") +
  theme(panel.background = element_rect(fill = "gray98"),
        legend.position="right", legend.background = element_rect(fill = alpha("white", 0.5), color = "gray"))
```



```{r}
otp_stop()
```
