---
title: "Assignment-7_v1"
author: "Gemma Holt"
date: "10/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```


## Setting up
```{r}
library(tidyverse)
library(sf)
library(leaflet)
library(htmltools)
library(htmlwidgets)
library(raster)
library(gstat)
library(spatial)
library(ggthemes)
library(ggspatial)
library(viridis)
```


## Loading the data
Inspired by a conversation with my neighbor Betty, I was curious if there was public data about rat sightings. (Betty and I spent a lot of time kvetching about the neighborhood rats.) Turns out, there is a database of 311 calls! While the premise of this is a *little* silly, I do think it could be useful as a mechanism of tracking the changes in the rat population in Somerville over time.
```{r}
somerville_plane <- "+proj=lcc +lat_1=41.71666666666667 +lat_2=42.68333333333333 +lat_0=41 +lon_0=-71.5 +x_0=200000 +y_0=750000 +ellps=GRS80 +units=m +no_defs"

somerville_rats <- read_csv("C:/Users/gemma/Desktop/gemmaholt-vis/Assignment-7-files/Rat_Data.csv") %>%
  mutate_at(vars(lat, lon), as.numeric) %>%
  st_as_sf(coords=c("lat","lon"), agr = "constant", crs=somerville_plane, stringsAsFactors = FALSE, remove = TRUE)

somerville_nhoods <- st_read("C:/Users/gemma/Desktop/gemmaholt-vis/Assignment-7-files/Neighborhoods/Neighborhoods.shp", quiet = TRUE) %>%
  dplyr::select(NBHD) %>%
  st_transform(somerville_plane)
```


## Mapping Somerville Neighborhoods
```{r}
leaflet(somerville_nhoods) %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addPolygons(highlightOptions = highlightOptions(fillColor = "yellow",
                                                  fillOpacity = 1),
              label = ~NBHD,
              weight = 1)
```


## Combining datasets
```{r}
nhood_rats <- st_join(somerville_rats, somerville_nhoods, join = st_within)
```

