---
title: "Assignment 7 Final"
author: "Gemma Holt"
date: "10/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```


## Setting up
```{r}
library(tidyverse)
library(sf)
library(leaflet)
library(htmltools)
library(htmlwidgets)
library(raster)
library(gstat)
library(spatial)
library(ggthemes)
library(ggspatial)
library(viridis)
```


## Loading the data
Inspired by a conversation with my neighbor Betty, I was curious if there was public data about rat sightings. (Betty and I spent a lot of time kvetching about the neighborhood rats.) Turns out, there is a database of 311 calls! While the premise of this is a *little* silly, I do think it could be useful as a mechanism of tracking the changes in the rat population in Somerville over time.
```{r}
coordinate_plane <- "+proj=longlat +datum=WGS84"

somerville_rats <- read_csv("C:/Users/gemma/Desktop/gemmaholt-vis/Assignment-7-files/Rat_Data.csv") %>%
  mutate_at(vars(lat, lon), as.numeric) %>%
  drop_na()

somerville_nhoods <- st_read("C:/Users/gemma/Desktop/gemmaholt-vis/Assignment-7-files/Neighborhoods/Neighborhoods.shp", quiet = TRUE) %>%
  dplyr::select(NBHD) %>%
  st_transform(coordinate_plane)
```

## Combining datasets
This is basically a joke. 
```{r}
rat_icon <- makeIcon(
  iconUrl = "https://emojipedia-us.s3.dualstack.us-west-1.amazonaws.com/thumbs/240/apple/237/rat_1f400.png",
  iconWidth = 38, iconHeight = 38)

leaflet(somerville_nhoods) %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addPolygons(highlightOptions = highlightOptions(fillColor = "yellow",
                                                  fillOpacity = 1),
              label = ~NBHD,
              weight = 1) %>%
  addMarkers(lng = somerville_rats$lon,
             lat= somerville_rats$lat, icon = rat_icon)


```

## Place-based rats
Turning rats into spatial objects. 
```{r}
somerville_rats_sf <- st_as_sf(somerville_rats, coords=c("lon","lat")) %>%
  st_set_crs(4326) %>%
  st_transform(coordinate_plane)

nhood_rats <- somerville_nhoods %>%
mutate(num_rats = lengths(st_contains(somerville_nhoods, somerville_rats_sf)))
```


## Analyzing the distribution of rats
```{r}
nhood_rats$label <-
  paste(nhood_rats$NBHD, "<br>",
        nhood_rats$num_rats, "rat incidents") %>%
  lapply(htmltools::HTML)

bins <- seq(min(nhood_rats$num_rats),
            max(nhood_rats$num_rats), by = 1)
pal <- colorNumeric("viridis", 
                    domain = nhood_rats$num_rats,
                    na.color = "#00000000")

leaflet(nhood_rats) %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addPolygons(highlightOptions = highlightOptions(fillOpacity = 1),
              label = ~label,
              fillColor = ~pal(num_rats),
              weight = 1, color = "black") %>% 
  addLegend(pal = pal, 
            values = ~num_rats,
            bins = 3,
            opacity = .7, title = "Number of rat incidents",
            position = "topright")
```


## Rats as points
```{r}
MA_state_plane <- "+proj=lcc +lat_1=41.71666666666667 +lat_2=42.68333333333333 +lat_0=41 +lon_0=-71.5 +x_0=200000 +y_0=750000 +ellps=GRS80 +units=m +no_defs "

WGS84 <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"

nhood_points <- st_centroid(
  st_transform(nhood_rats, crs = MA_state_plane)) %>%
  st_transform(WGS84)

rat_icon <- makeIcon(
  iconUrl = "https://emojipedia-us.s3.dualstack.us-west-1.amazonaws.com/thumbs/240/apple/237/rat_1f400.png",
  iconWidth = 38, iconHeight = 38)

leaflet(nhood_points) %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addMarkers(label = ~label,
             icon = rat_icon) %>% 
  addControl("Number of rat incidents",
            position = "topright")
```



## Rat-ster map
(I'm sorry.) 
```{r}
nhood_pts_sp <- nhood_points %>%
  st_transform(MA_state_plane) %>%
  as_Spatial()

nhood_poly_sp <- somerville_nhoods %>%
  st_transform(MA_state_plane) %>%
  as_Spatial()

somerville_raster <- raster(nhood_poly_sp, res=10)

gs <- gstat(formula=num_rats~1, locations=nhood_pts_sp)
idw_interp <- interpolate(somerville_raster, gs)

idw_interp_clip <- mask(idw_interp, nhood_poly_sp)

leaflet(nhood_points) %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addRasterImage(idw_interp_clip, colors = pal, opacity = 0.8) %>% 
  addLegend(pal = pal, 
            values = ~num_rats,
            bins = 3,
            opacity = 0.7, title = "Estimated rat incidents",
            position = "topright")

```


## Discussion
Overall, I think the most informative map is the chloropleth map. As shown, it indicates neighborhood trends in a way that is coherent and easy to understand (unlike, for example, the rat icon map, however enjoyable I find it.) This is also probably the most informative map, particularly if you were trying to manage the population in a spatially targeted way. 

However, I do think that this data lends itself well to interpolation and that the raster map is the most interesting. It may not be an accurate conclusion, but the raster map does seem to indicate that the rat population in Somerville radiates out from a large colony in Union Square. I don't know whether this is accurate as I am basing my understanding of rat populations on the Pixar movie Ratatouille. In my introduction, I hypothesized that interpolation could be useful as a mechanism for tracking changes in population over time. I still think that this holds, assuming that 311 reports are an accurate proxy. 