---
title: "Assignment-7_v1"
author: "Gemma Holt"
date: "10/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```


## Setting up
```{r}
library(tidyverse)
library(sf)
library(leaflet)
library(htmltools)
library(htmlwidgets)
library(raster)
library(gstat)
library(spatial)
library(ggthemes)
library(ggspatial)
library(viridis)
```


## Loading the data
Inspired by a conversation with my neighbor Betty, I was curious if there was public data about rat sightings. (Betty and I spent a lot of time kvetching about the neighborhood rats.) Turns out, there is a database of 311 calls! While the premise of this is a *little* silly, I do think it could be useful as a mechanism of tracking the changes in the rat population in Somerville over time.
```{r}
coordinate_plane <- "+proj=longlat +datum=WGS84"

somerville_rats <- read_csv("C:/Users/gemma/Desktop/gemmaholt-vis/Assignment-7-files/Rat_Data.csv") %>%
  mutate_at(vars(lat, lon), as.numeric) %>%
  drop_na()

somerville_nhoods <- st_read("C:/Users/gemma/Desktop/gemmaholt-vis/Assignment-7-files/Neighborhoods/Neighborhoods.shp", quiet = TRUE) %>%
  dplyr::select(NBHD) %>%
  st_transform(coordinate_plane)
```


## Mapping Somerville Neighborhoods
```{r}
leaflet(somerville_nhoods) %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addPolygons(highlightOptions = highlightOptions(fillColor = "yellow",
                                                  fillOpacity = 1),
              label = ~NBHD,
              weight = 1)
```


## Combining datasets
This is basically a joke. 
```{r}
rat_icon <- makeIcon(
  iconUrl = "https://emojipedia-us.s3.dualstack.us-west-1.amazonaws.com/thumbs/240/apple/237/rat_1f400.png",
  iconWidth = 38, iconHeight = 38)

leaflet(somerville_nhoods) %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addPolygons(highlightOptions = highlightOptions(fillColor = "yellow",
                                                  fillOpacity = 1),
              label = ~NBHD,
              weight = 1) %>%
  addMarkers(lng = somerville_rats$lon,
             lat= somerville_rats$lat, icon = rat_icon)


```

## Place-based rats
Turning rats into spatial objects. 
```{r}
somerville_rats_sf <- st_as_sf(somerville_rats, coords=c("lon","lat")) %>%
  st_set_crs(4326) %>%
  st_transform(coordinate_plane)

nhood_rats <- somerville_nhoods %>%
mutate(num_rats = lengths(st_contains(somerville_nhoods, somerville_rats_sf)))
```


## Analyzing the distribution of rats
```{r}
nhood_rats$label <-
  paste(nhood_rats$NBHD, "<br>",
        nhood_rats$num_rats, "rat incidents") %>%
  lapply(htmltools::HTML)

bins <- seq(min(nhood_rats$num_rats),
            max(nhood_rats$num_rats), by = 1)
pal <- colorNumeric("viridis", 
                    domain = nhood_rats$num_rats,
                    na.color = "#00000000")

leaflet(nhood_rats) %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addPolygons(highlightOptions = highlightOptions(fillOpacity = 1),
              label = ~label,
              fillColor = ~pal(num_rats),
              weight = 1, color = "black") %>% 
  addLegend(pal = pal, 
            values = ~num_rats,
            bins = 3,
            opacity = .7, title = "Number of rat incidents",
            position = "topright")
```


## Rats as points
```{r}
MA_state_plane <- "+proj=lcc +lat_1=41.71666666666667 +lat_2=42.68333333333333 +lat_0=41 +lon_0=-71.5 +x_0=200000 +y_0=750000 +ellps=GRS80 +units=m +no_defs "

WGS84 <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"

nhood_points <- st_centroid(
  st_transform(nhood_rats, crs = MA_state_plane)) %>%
  st_transform(WGS84)

rat_icon <- makeIcon(
  iconUrl = "https://emojipedia-us.s3.dualstack.us-west-1.amazonaws.com/thumbs/240/apple/237/rat_1f400.png",
  iconWidth = 38, iconHeight = 38)

leaflet(nhood_points) %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addMarkers(label = ~label,
             icon = rat_icon) %>% 
  addControl("Number of rat incidents",
            position = "topright")
```



## Rat-ster map
(I'm sorry.) 
```{r}
nhood_pts_sp <- nhood_points %>%
  st_transform(MA_state_plane) %>%
  as_Spatial()

nhood_poly_sp <- somerville_nhoods %>%
  st_transform(MA_state_plane) %>%
  as_Spatial()

somerville_raster <- raster(nhood_poly_sp, res=10)

gs <- gstat(formula=num_rats~1, locations=nhood_pts_sp)
idw_interp <- interpolate(somerville_raster, gs)

idw_interp_clip <- mask(idw_interp, nhood_poly_sp)

leaflet(nhood_points) %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addRasterImage(idw_interp_clip, colors = pal, opacity = 0.8) %>% 
  addLegend(pal = pal, 
            values = ~num_rats,
            bins = 3,
            opacity = 0.7, title = "Estimated rat incidents",
            position = "topright")

```

