---
title: "Assignment 2"
author: "Gemma Holt"
date: "9/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Projections and Distortions

## Setup
```{r}
library(tidyverse)
library(ggplot2)
library(ggthemes)
library (rnaturalearth)
library(cartogram)
library(sf)
library(ggspatial)
library(rnaturalearthdata)
library(viridis)
library(dplyr)
```

## Finland Base Map
```{r}
finland_regions <- ne_states(country= "Finland", returnclass= "sf")

finland_map <- ggplot(finland_regions) +
  geom_sf()+
  theme_bw()

finland_map
```

##Projections
```{r}
finn_UCS <- "+proj=tmerc +lat_0=0 +lon_0=27 +k=1 +x_0=3500000 +y_0=0 +ellps=intl +units=m +no_defs"

finn_zone0 <- "+proj=tmerc +lat_0=0 +lon_0=18 +k=1 +x_0=500000 +y_0=0 +ellps=intl +units=m +no_defs"

finn_zone5 <- "+proj=tmerc +lat_0=0 +lon_0=33 +k=1 +x_0=5500000 +y_0=0 +ellps=intl +units=m +no_defs"
```


## Finland Uniform Coordinate System
```{r}
ggplot(finland_regions)+
  geom_sf() +
  coord_sf(crs = finn_UCS) +
  theme_bw()+
xlab("Longitude") + ylab("Latitude") +
    ggtitle("Finland")
```

## Finland Zone 0
This zone is designed for mapping the western-most (Sweden-adjacent) part of Finland.
```{r}
ggplot(finland_regions)+
  geom_sf() +
  coord_sf(crs = finn_zone0) +
  theme_bw()+
  xlab("Longitude") + ylab("Latitude") +
    ggtitle("Finland", subtitle="Zone 0")
```


## Finland Zone 5
This zone is designed for mapping the eastern-most (Russia-adjacent) part of Finland. 
```{r}
ggplot(finland_regions)+
  geom_sf() +
  coord_sf(crs = finn_zone5) +
  theme_bw()+
  xlab("Longitude") + ylab("Latitude") +
    ggtitle("Finland", subtitle="Zone 5")
```

## Transforming coordinate systems
```{r}
regions_transformed <- finland_regions %>%
  st_transform(crs = finn_UCS)

ggplot(regions_transformed) +
  geom_sf(color = "black", fill = "blue") +
  theme_bw()
```


## Importing data
I downloaded regional population and immigration data from Statistics Finland. I ran into a problem since the csv file was not exporting umlauts. I addressed most of the issues in the csv, excluding the data for 	
Päijät-Häme which I changed manually. 
```{r}
finn_migrate <- read_csv("C:/Users/gemma/OneDrive/GSD Semester 1/Spatial Analysis/Gemma_Holt_VIS/Github/gemmaholt-vis/Assignment-2_files/FinlandImmigration2.csv")

finland_skinny <- finland_regions %>% 
  select(name, latitude, longitude, geometry)
view(finland_skinny)

finland_merge <- left_join(finland_skinny,finn_migrate, by="name")
View(finland_merge)

finland_merge[18,4] = 199604
finland_merge[18,5] = 11135

finland_merge <- mutate(finland_merge, pct_migrate=100*var_migrate/var_pop)

finland_merge <- mutate(finland_merge,migrate_10k=var_migrate/1000)

##transform(finland_merge, "Persons born abroad" = as.numeric(as.character("Persons born abroad")))
```


## Doing some stuff
```{r}
ggplot(finland_merge, aes(fill = pct_migrate))+
geom_sf(color = NA) +
  scale_fill_viridis_c(
    name = "Percent",
    breaks = seq(0, 15, by = 2),
    labels = formatC(seq(0, 15, by = 2))) + 
  theme_map() +
  theme(legend.background = element_blank(),legend.position="right", 
        legend.title = element_text(size=9)) +
    ggtitle("Percent of population born abroad")
```


## Doing some other stuff
```{r}
lseq <- function(from=1, to=100000, length.out=6) {
  # logarithmic spaced sequence
  # blatantly stolen from library("emdbook"), because need only this
  # I blatantly stole this from https://stackoverflow.com/questions/23901907/create-a-log-sequence-across-multiple-orders-of-magnitude
  exp(seq(log(from), log(to), length.out = length.out))
}

ggplot(finland_merge, aes(fill = migrate_10k))+
geom_sf(color = NA) +
  scale_fill_viridis_c(trans = "log",
    name = "Number",
    breaks = lseq(1, 250, 10),
    labels = formatC(lseq(1, 250, 10))) + 
  theme_map() +
  theme(legend.background = element_blank(),legend.position="right", 
        legend.title = element_text(size=9)) +
    ggtitle("Number of persons born abroad per 10,000")
```


## Doing some more stuff
Code inspired by Carole's density equalized plot
```{r}
library(cartogram)
migrate_cartogram_cont <- finland_merge %>%
  st_transform(crs = finn_UCS) %>%
  cartogram_cont("var_migrate")
```


```{r}
ggplot(migrate_cartogram_cont, aes(fill = var_migrate)) +
  geom_sf(color = NA) +
  scale_fill_viridis_c(
    name = "Number of persons born abroad",
    breaks = seq(0, 250000, by = 25000),
    labels = formatC(seq(0, 250000, by = 25000), 
                     big.mark = ",", format = "f", digits = 0)) +
  theme_map() +
  theme(legend.background = element_blank(),legend.position="right", 
        legend.title = element_text(size=9)) +
    ggtitle("Number of persons born abroad")
```