---
title: "Aspatial Data Analysis"
author: "Gemma Holt"
date: "9/9/2020"
output:
  html_document:
    theme: lumen
    toc: yes
    toc_depth: 3   
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, error = FALSE)
```

# Relationships and Distributions


## Setup
```{r}
library(tidycensus)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(ggthemes)
library(wesanderson)
library(viridis)
library(ggExtra)
library(extrafont)
library(RColorBrewer)
library(extrafontdb)
```


## Loading the data

These variables were copied from Carole's example problem set. I used data from King County, which is where I grew up. Additionally, after class on Monday I added an additional variable (inspired by Miguel) in order to convert the percentage of rental households into a categorical variable.  

```{r}
vars <- c(tot_hh = "H001001",
         tot_rural_hh = "H002005",
         tot_rent_hh = "H004004",
         tot_own_hh = "H004002",
         avg_hh_size = "H012001",
         tot_pop = "P003001",
         nh_white_pop = "P005003",
         nh_black_pop = "P005004",
         nh_asian_pop = "P005006",
         hispanic_pop = "P005010",
         med_age = "P013001")

KC_tracts <- get_decennial(geography = "tract", variables = vars,state = "WA", county = "King", output = "wide") %>%

  mutate(rural = (tot_rural_hh == tot_hh) ,
         pct_rental = tot_rent_hh / tot_hh,
         pct_own = tot_own_hh / tot_hh,
         maj_house = case_when(
            tot_own_hh / tot_hh > 0.5 ~ "owned",
            tot_rent_hh / tot_hh > 0.5 ~ "rented",
            TRUE ~ "none"),
         majority_race_eth = case_when(
           nh_white_pop / tot_pop > 0.5 ~ "nh_white",
           nh_black_pop / tot_pop > 0.5 ~ "nh_black",
           nh_asian_pop / tot_pop > 0.5 ~ "nh_asian",
           hispanic_pop / tot_pop > 0.5 ~ "hispanic",
           (nh_white_pop + nh_black_pop + nh_asian_pop + hispanic_pop) /
             tot_pop < 0.5 ~ "other",
           TRUE ~ "None")) %>%
  filter(tot_pop > 0, tot_hh > 0) %>%
  select(GEOID, rural, maj_house, majority_race_eth, 
         pct_rental, avg_hh_size, tot_pop, med_age)

```


## Graphs and Figures
### Scatterplot
```{r}
scatterplot <- ggplot(KC_tracts, aes(x = pct_rental,
           y = avg_hh_size,)) +
  geom_point() +
  geom_smooth(model=lm) +
  scale_x_continuous(name = "Percent of renter occupied units",
      breaks = seq(0, 1, by = 0.1),
      labels = paste(seq(0, 100, by = 10), "%", sep = "")) +
  scale_y_continuous(name = "Average Household Size",
                     breaks = seq(0, 6, by = .5),
                     labels = paste(seq(0, 6, by =.5), sep = "")) +
  labs(title="Scatter Plot",
       subtitle="Average household size vs percent of renter occupied units")+
theme_few()+
  theme(text = element_text(family = "Verdana"))

scatterplot

png(file="C:/Users/gemma/OneDrive/GSD Semester 1/Spatial Analysis/Final Portfolio/Figures/scatterplot.png",
width=600, height=350)
scatterplot
dev.off()
```

### Violin Plot
```{r}
violinplot <- ggplot(KC_tracts, aes(x=majority_race_eth, y=avg_hh_size, fill = majority_race_eth)) + 
  geom_violin() +
  labs(title="Violin Plot",
       subtitle="Average household size vs majority race/ethnicity",
       x="Majority race/ethnicity",
       y="Average Household Size")+
  scale_fill_manual(values = brewer.pal(n=3, name="YlGnBu")) +
  scale_x_discrete(name = "Majority race/ethnicity",
                   labels = c("Non-Hispanic Asian",
                                  "Non-Hispanic white",
                                  "No majority"))+
theme_few()+
theme(legend.position = "none")+
theme(text = element_text(family = "Verdana"))

violinplot

png(file="C:/Users/gemma/OneDrive/GSD Semester 1/Spatial Analysis/Final Portfolio/Figures/violinplot.png",
width=600, height=350)
violinplot
dev.off()
                      
```

### Stacked Bar Graph

This figure was also inspired by Miguel. I had originally used a histogram to show the distribution of rental households by majority race/ethnicity, but because over 75% of tracts are majority white, the data was not very interesting. Using 100% stacked bars, you can more clearly see the differences. Thanks Miguel!
```{r}
stackedbar <- ggplot(KC_tracts, aes(x = majority_race_eth, fill = maj_house)) +
  geom_bar(position = "fill") +
  scale_x_discrete(name = "Majority race/ethnicity",
                       labels = c("Non-Hispanic Asian",
                                  "Non-Hispanic white",
                                  "No majority")) +
  scale_y_continuous(name = "Percent of Tracts",
                     breaks = seq (0,1, by = 0.1),
                     labels = paste(seq(0, 100, by=10), "%", sep = ""))+
scale_fill_manual(name = "Rented or owned",
                  values = brewer.pal(n=3, name="YlGnBu"), 
                  labels = c("Majority of households owned", "Majority of households rented")) +
  labs(title="Bar Plot",
       subtitle="Majority race/ethnicity by homeownership status")+
  theme_few()+
theme(text = element_text(family = "Verdana"))

stackedbar

png(file="C:/Users/gemma/OneDrive/GSD Semester 1/Spatial Analysis/Final Portfolio/Figures/stackedbar.png",
width=600, height=350)
stackedbar
dev.off()
```

### Density Plot
```{r}
densityplot <- ggplot(KC_tracts, aes(x = pct_rental, y=med_age)) +
  stat_density2d(aes(fill = stat(level)), geom="polygon") + scale_fill_viridis(option = "plasma") +
    labs(x = "Percent of renter occupied units",
       y = "Median Age", 
       title = "Density Plot",
       subtitle = "Median age vs percent of renter occupied units")+
  theme_few()+
  theme(text = element_text(family = "Verdana"))

densityplot

png(file="C:/Users/gemma/OneDrive/GSD Semester 1/Spatial Analysis/Final Portfolio/Figures/densityplot.png",
width=600, height=350)
densityplot
dev.off()
```
 
 
###  Area Plot
```{r}
 areaplot <- ggplot(KC_tracts, aes(x=avg_hh_size)) +
  geom_density(fill="#2C7FB8", color="#7FCDBB", alpha=0.8) +
  scale_x_continuous(name = "Average Household Size",
                     breaks = seq(0, 6, by = .5),
                    labels = paste(seq(0, 6, by = .5), sep = "")) +
  labs(title="Area Plot",
       subtitle="Average Household Size in King County")+
  theme_few()+
  theme(text = element_text(family = "Verdana"))

areaplot

png(file="C:/Users/gemma/OneDrive/GSD Semester 1/Spatial Analysis/Final Portfolio/Figures/areaplot.png",
width=600, height=350)
areaplot
dev.off()
```

### Histogram
```{r}
densityhistplot <- ggplot(KC_tracts, aes(x=med_age)) +
  geom_histogram(aes(y=..density..),color="black", fill="white") +
  geom_density(alpha=.2, fill="#2C7FB8") +
  labs(title="Density Plot with Histogram",
       subtitle="Median Age in King County")+
  theme_few()+
  theme(text = element_text(family = "Verdana"))

densityhistplot

png(file="C:/Users/gemma/OneDrive/GSD Semester 1/Spatial Analysis/Final Portfolio/Figures/densityhistplot.png",
width=600, height=350)
densityhistplot
dev.off()
```

### Scatterplot with marginal histogram
```{r}
g <- ggplot(KC_tracts, aes(x = med_age, y= avg_hh_size, color = maj_house)) +
  geom_point() +
  scale_x_continuous(name = "Median Age",
      breaks = seq(0, 100, by = 10)) +
  scale_y_continuous(name = "Average Household Size",
                     breaks = seq(0, 6, by = .5),
                     labels = paste(seq(0, 6, by =.5), sep = "")) +
  scale_color_manual(name = "Rented or owned",
                     values = brewer.pal(n=3, name="YlGnBu"), 
                     labels = c("Majority of households owned", "Majority of households rented")) +
  labs(title="Scatter Plot with Marginal Histogram",
       subtitle="Average household size vs median age")+
  theme_few()+
  theme(legend.position="bottom")+
  theme(text = element_text(family = "Verdana"))

marginalplot <- ggMarginal(g, type = "histogram", fill = "#1D91C0")

marginalplot 

png(file="C:/Users/gemma/OneDrive/GSD Semester 1/Spatial Analysis/Final Portfolio/Figures/marginalplot.png",
width=600, height=350)
marginalplot
dev.off()

```

### Pie Chart
```{r}
pie_chart <- ggplot(KC_tracts, aes(x = 1, fill = majority_race_eth)) +
  geom_bar(position = "fill") +
  coord_polar(theta = "y") +
  scale_y_continuous(name = "Number of tracts") +
  scale_fill_manual(values = brewer.pal(n=3, name="YlGnBu"), name = "Majority race/ethnicity", labels = c("Non-Hispanic\nAsian",
                                  "Non-Hispanic\nwhite",
                                  "No majority")) +
  theme_void() +
  labs(title ="Pie Chart",
       subtitle ="Majority Race/Ethnicity in King County")+
  theme(text = element_text(family = "Verdana"))

pie_chart

png(file="C:/Users/gemma/OneDrive/GSD Semester 1/Spatial Analysis/Final Portfolio/Figures/pie_chart.png",
width=600, height=350)
pie_chart
dev.off()

```

### Heat Map
```{r}
 heatmap <- ggplot(KC_tracts, aes(x = rural, y = majority_race_eth, fill = pct_rental)) +
       geom_tile() +
      scale_y_discrete(name = "Majority race/ethnicity",
                       labels = c("Non-Hispanic\nAsian",
                                  "Non-Hispanic\nwhite",
                                  "No majority")) +
      scale_x_discrete(name = "Majority income",
                       labels = c("Urban tract",
                                  "Rural Tract")) +
      scale_fill_distiller(name = "Percent Rental",
                           palette = "YlGrBu") +
      labs(title ="Heat Map")+ 
      theme_few()

heatmap

png(file="C:/Users/gemma/OneDrive/GSD Semester 1/Spatial Analysis/Final Portfolio/Figures/heatmap.png",
width=600, height=350)
heatmap
dev.off()
```

### Histogram
```{r}
histogram <- ggplot(KC_tracts, aes(pct_rental)) + scale_fill_brewer(name="Majority race/ethnicity",                          palette = "YlGrBu",
                  labels = c("Non-Hispanic\nAsian",
                                  "Non-Hispanic\nwhite",
                                  "No majority")) + 
  geom_histogram(aes(fill=majority_race_eth), 
                   bins=5, 
                   col="black", 
                   size=.1) +
  scale_x_continuous(name = "Percent rental") +
  labs(title="Histogram with Fixed Bins", 
       subtitle="Percent of renter occupied units by majority race/ethnicity")+
  theme_few()+
  theme(text = element_text(family = "Verdana"))

histogram

png(file="C:/Users/gemma/OneDrive/GSD Semester 1/Spatial Analysis/Final Portfolio/Figures/histogram.png",
width=600, height=350)
histogram
dev.off()
```
 
 